---
title: "Tidy Text"
output: html_document
date: "2025-10-15"
---

# Exercise 1
Load the `tidyverse` package. Install and load the `tidytext` and `topicmodels` packages.
```{r}
# Load packages here
library(tidyverse)
library(tidytext)
library(topicmodels)

```

Run this code, so that `jss` is in your global environment.
```{r}
data("JSS_papers")
jss <- JSS_papers |> 
  as_tibble() |> 
  mutate(date = ymd(date),
         year = year(date),
         abstract = unlist(description)) |> 
  select(year, abstract)

View(jss)
```


# Exercise 2
Tidy `jss` so that there is one token (word) per row.
```{r}
jss_tokens <- jss %>%
  # Your code goes here
  unnest_tokens(word, abstract) 

```

# Exercise 3
Remove all stop words from `jss_tokens`, that are included in the SMART lexicon.
```{r}
jss_tokens_keywords <- jss_tokens %>%
  # Your code goes here
  anti_join(filter(stop_words, lexicon == "SMART"))
  
```

# Exercise 4
What are the five most common words in the most recent year of data?
```{r}
# five most common, most recent year
jss_tokens_keywords |> 
  # your code goes here
  filter(year == max(year)) |> 
  count(word, sort = TRUE) |> 
  slice(1:5)
  
```

What are the five most common words in the earliest year of data?
```{r}
# five most common, earliest year
jss_tokens_keywords |> 
  # your code goes here
  filter(year == min(year)) |> 
  count(word, sort = TRUE) |> 
  slice(1:5)

```

# Exercise 5
Run the code below to see how many times "r" (as in the "R" programming language) appears in `jss_tokens_keywords`.

Do you think that frequency value represents the true number of times "R" is mentioned in JSS abstracts between 1996-2010? If not, what might be causing the discrepancy? **The letter `r` is in the SMART lexicon.**

```{r}
# Get R frequency
jss_tokens_keywords |> 
  filter(word == "r") |> 
  nrow()

```

# Exercise 6
Using `jss_tokens`, make a line plot with `year` on the x-axis and the yearly frequency for the token `r` on the y-axis. (Hint: use `filter()` and then `count()`.)

What do you notice looking at the plot?

```{r}
jss_tokens |> 
  # Your code goes here
  filter(word == "r") |> 
  count(year) |> 
  ggplot(aes(x = year, y = n)) +
  geom_line()

```

